\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}  
I would like to express my deepest gratitude to those who have supported me
during this academic journey.

First and foremost, I extend my profound thanks to my supervisor, Marco, and
co-supervisor, Carlo. Their invaluable support throughout the research process
and the writing of this dissertation has been crucial. From the first day of
working on this thesis, you have shown me great kindness and were exceptionally
accessible for all matters, extending even beyond academics. If working on the
thesis has been a very enjoyable experience for me, it is largely thanks to you.

I must also extend a heartfelt thank you to my parents. Their steadfast support
and belief in my abilities have afforded me the opportunity to pursue my
studies. Their love and patience, especially during the most challenging times,
have been my foundation. Without their sacrifices and encouragement, none of
this would have been possible.

My journey at the university was also enriched by peers who became close
friends. To Edoardo and Giuseppe, with whom I've shared the long and memorable
journey of university life, thank you. The unforgettable moments we shared truly
brightened my days.

Finally, I would like to thank my girlfriend, Cecilia. Even though you joined me
during the final stretch of this beautiful journey, you have proven to be an
invaluable partner, both patient and always ready to listen. I also extend my
gratitude to all other friends who have been part of this journey, whose
companionship and support have been invaluable. Thank you all for being part of
this story.